---
title: "Data pre-validation"
output: html_document
---

## 1. Satellite Data

GEE data pre-validation code link: <https://code.earthengine.google.com/5af2353146d900b5380b4e6f28f78b04>

More details in Word document saved in OneDrive.

## 2. GTFS Data

### r5r Preparation

**Install JDK using {rJavaEnv} package:**

```         
Downloaded in: C:/Users/choic/AppData/Local/R/cache/R/rJavaEnv
```

```{r}
# install.packages('rJavaEnv')

# check version of Java currently installed (if any) 
rJavaEnv::java_check_version_rjava()

# install Java 21
rJavaEnv::java_quick_install(version = 21)
```

**Install r5r:**

```{r}
# Install from CRAN
# install.packages("r5r")

# Warning message: Currently, Java memory is set to "-Xmx512m"
# Warning message suggest set Java max memory to 2 GB before loading r5r
options(java.parameters = '-Xmx2G')
# ChatGPT suggested `options(java.parameters = "-Xmx4G")`
library(r5r)
```

**Data requirements for r5r :** <https://ipeagit.github.io/r5r/#data-requirements>

### 2.1. London

#### 2.1.1. Data Sources

+------------------------+----------------------------------------------------+-----------------------------------------------------------------------------------+
| Data                   | Description                                        | Link                                                                              |
+========================+====================================================+===================================================================================+
| GTFS                   | Download as NaPTAN file                            | <https://beta-naptan.dft.gov.uk/>                                                 |
|                        |                                                    |                                                                                   |
|                        | It then needs to be converted into a GTFS format.  |                                                                                   |
+------------------------+----------------------------------------------------+-----------------------------------------------------------------------------------+
| Road Network from OSM  | Downloaded as .pbf file                            | <https://download.geofabrik.de/europe/united-kingdom/england/greater-london.html> |
+------------------------+----------------------------------------------------+-----------------------------------------------------------------------------------+

#### 2.1.2. Convert data into GTFS format

```{r}
remotes::install_github("ITSleeds/UK2GTFS")
```

```{r}

```

### 2.2. New York

#### 2.1.2. Data Sources
