---
title: "London Data pre-validation"
output: html_document
---

## 1. Satellite Data

GEE data pre-validation code link: <https://code.earthengine.google.com/5af2353146d900b5380b4e6f28f78b04>

**London Data Availability**

GEE code link: <https://code.earthengine.google.com/ebd45ea5ac18517098795c096c8411d8>

| Year | Landsat 8 | Landsat 9 |
|------|-----------|-----------|
| 2022 | 3 images  | 2 images  |
| 2024 | 0 image   | 1 image   |
| 2025 | 0 image   | 2 images  |

## 2. GTFS Data

### r5r Preparation

**Install JDK using {rJavaEnv} package:**

```         
Downloaded in: C:/Users/choic/AppData/Local/R/cache/R/rJavaEnv
```

```{r}
# install.packages('rJavaEnv')

# check version of Java currently installed (if any) 
rJavaEnv::java_check_version_rjava()

# install Java 21
rJavaEnv::java_quick_install(version = 21)
```

**Install r5r:**

```{r}
# Install from CRAN
# install.packages("r5r")

# Warning message: Currently, Java memory is set to "-Xmx512m"
# Warning message suggest set Java max memory to 2 GB before loading r5r
options(java.parameters = '-Xmx2G')
# ChatGPT suggested `options(java.parameters = "-Xmx4G")`
library(r5r)
```

**Data requirements for r5r :** <https://ipeagit.github.io/r5r/#data-requirements>

### 2.1. London

#### 2.1.1. Data Sources

+-----------------------+---------------------------------------------------------------------------------------------------------------------------------------+----------------------------------------------------------------------------------------+
| Data                  | Description                                                                                                                           | Link                                                                                   |
+=======================+=======================================================================================================================================+========================================================================================+
| GTFS                  | Download as NaPTAN file                                                                                                               | <https://beta-naptan.dft.gov.uk/>                                                      |
|                       |                                                                                                                                       |                                                                                        |
|                       | It then needs to be converted into a GTFS format (<https://citygeographics.org/r5r-workshop/uk-transit-data-transxchange-and-atoc/>). |                                                                                        |
+-----------------------+---------------------------------------------------------------------------------------------------------------------------------------+----------------------------------------------------------------------------------------+
| GTFS                  | Download L.zip (for London)                                                                                                           | Look for Core FTP and use the access host, username and password from the sent email.  |
+-----------------------+---------------------------------------------------------------------------------------------------------------------------------------+----------------------------------------------------------------------------------------+
| Road Network from OSM | Downloaded as .pbf file                                                                                                               | <https://download.geofabrik.de/europe/united-kingdom/england/greater-london.html>      |
+-----------------------+---------------------------------------------------------------------------------------------------------------------------------------+----------------------------------------------------------------------------------------+

#### 2.1.2. Convert data into GTFS format

```{r}
remotes::install_github("ITSleeds/UK2GTFS")
```

```{r}
# Convert London TNDS (TransXChange) data into a GTFS object for later use in r5r

# Load the specific package for converting to GTFS
library(UK2GTFS)

# Load Packages
library(readr)
library(dplyr)


# Read NaPTAN stop data and rename columns for use with UK2GTFS
naptan_local <- read_csv("r5r_london/Stops.csv", show_col_types = FALSE) %>%
  rename(
    stop_id = ATCOCode,
    stop_name = CommonName,
    stop_lat = Latitude,
    stop_lon = Longitude,
    stop_type = StopType
  )

# %>%
#    filter(!is.na(stop_id),
#          !is.na(stop_name),
#          !is.na(stop_lat),
#          !is.na(stop_lon))



# Using the L.zip (London specific file) and convert into a GTFS object
path_in <- "r5r_london/L.zip"
gtfs <- transxchange2gtfs(path_in = path_in,
                          # remove this line if it is not local
                          # naptan = naptan_local,
                          ncores = 5)



# Using a London file, so this is not needed? 
# Oxford_LAs_wgs84 <- Oxford_LAs %>%
# st_transform(., 4326)
  
#gtfs_london <- gtfs_clip(gtfs, Oxford_LAs_wgs84)
```

```{r}
# Save as GTFS file
gtfs_write(gtfs, 
           folder = "r5r_london/",
           name = "gtfs_london")
```

### 2.2. New York

#### 2.1.2. Data Sources
